### 2015-08-03 motivation.tex
Changes from Greg.
tweaks to existing practice discussion

Packaging problems arise in HPC because a supercomputer's hardware, OS, and
filesystem are shared by many users with different requirements.  The classic
solution to this problem is to use virtual
machines (VMs)~\cite{barham2003xen,rosenblum1999vmware,smith2005architecture}
or lightweight virtualization techniques like Linux
containers~\cite{felter2014updated,merkel2014docker}. This model allows each
 user to have a personalized environment with its own package manager, and it
has been extremely successful for servers at cloud data centers. VMs typically
have near-native compute performance but low-level HPC network drivers still
exhibit major performance issues. VMs are not well supported on many
non-Linux operating systems, an issue for the lightweight
kernels of bleeding-edge Blue Gene/Q and Cray machines.
Finally, each VM still uses a traditional package manager,
so running many configurations still requires a large number of VMs.
For facilities, this profusion of VMs is a security concern since it 
complicates mandatory patching. Also, managing  a large number of VM 
environments is tedious for users.

@@

Many HPC sites address this problem using {\it environment modules}, which
allow users to ``load'' and ``unload'' such settings dynamically using simple
commands. Environment modules emerged in 1991, and there are many implementations~\cite{dotkit,furlani+:lisa91,furlani+:lisa96,mclay:lmod,mclay:lmod-tutorial}.
The most advanced of these, Lmod~\cite{mclay:lmod,mclay:lmod-tutorial},
provides software hierarchies similar to the naming conventions in
Table~\ref{tab:naming-conventions}. This allows users to load a software stack
quickly if they know which one is required.

@@

The EasyBuild~\cite{hoste+:pyhpc12} tool is in production use at
the University of Ghent and the J\"ulich Supercomputing Center.
It allows multiple versions to be installed at once.  Rather
than setting {\tt RPATHs}, it auto-generates module files
to manage the environment, and it is closely coupled with
Lmod~\cite{geimer+:hust14}.  EasyBuild groups the compiler, MPI, FFT, and
BLAS libraries together in a {\it toolchain} that can be used by
package files. This provides some degree of composability and
separates compiler flags and MPI concerns from client packages.

>>>>

Packaging problems arise in HPC because a supercomputer's hardware, OS, and
file system are shared by many users with different requirements.  The classic
solution to this problem is to use virtual
machines (VMs)~\cite{barham2003xen,rosenblum1999vmware,smith2005architecture}
or lightweight virtualization techniques like Linux
containers~\cite{felter2014updated,merkel2014docker}. This model allows each
 user to have a personalized environment with its own package manager, and it
has been extremely successful for servers at cloud data centers. VMs typically
have near-native compute performance but low-level HPC network drivers still
exhibit major performance issues. VMs are not well supported on many
non-Linux operating systems, an issue for the lightweight
kernels of bleeding-edge Blue Gene/Q and Cray machines.
Finally, each VM still uses a traditional package manager,
so running many configurations still requires a large number of VMs.
For facilities, this profusion of VMs is a security concern since it
complicates mandatory patching. Also, managing  a large number of VM
environments is tedious for users.

@@

Many HPC sites address this problem using {\it environment modules}, which
allow users to ``load'' and ``unload'' such settings dynamically using simple
commands. Environment modules emerged in 1991, and there are many implementations~\cite{dotkit,furlani+:lisa91,furlani+:lisa96,mclay:lmod,mclay:lmod-tutorial}.
The most advanced of these, Lmod~\cite{mclay:lmod,mclay:lmod-tutorial},
provides software hierarchies that are similar to the naming conventions in
Table~\ref{tab:naming-conventions} and allow users to load a software stack
quickly if they know which one is required.

@@

The EasyBuild~\cite{hoste+:pyhpc12} tool is in production use at
the University of Ghent and the J\"ulich Supercomputing Center.
It allows multiple versions to be installed at once.  Rather
than setting {\tt RPATHs}, it auto-generates module files
to manage the environment, and it is closely coupled with
Lmod~\cite{geimer+:hust14}.  EasyBuild groups the compiler, MPI, FFT, and
BLAS libraries together in a {\it toolchain} that can be used by
package files. The grouping provides some composability and
separates compiler flags and MPI concerns from client packages.
